Thesis Title

### 2022-03-26 chapters/chapter_2.tex
finish acronym list

As mentioned in Chapter \ref{sssec:shotgun_metagenomics}, shotgun metagenomic (SMg) approaches have been a growing interest to deliver clinically relevant results without \textit{a priori} knowledge of what to expect from a particular clinical sample or patient. 
The capacity to detect all potential pathogens in a sample has great potential utility in the diagnosis of infectious disease. 
However, it is unclear how the variety of available methods impacts the end results.

In this publication SMg was applied to nine body fluid samples and one tissue sample from patients at the University Medical Center Groningen (UMCG) with varying degrees of contamination: one sample from peritoneal fluid, five from pus , two from synovial fluid of knees with prosthesis, one from sputum and one from a bone biopsy. The results of microbial identification through whole genome sequencing (WGS) and SMg were compared to standard culture-based microbiological methods.

In order to evaluate and compare the accuracy and reliability of the bioinformatics analyses in providing the closest results to culture and WGS of any cultured isolates, three different bioinformatic pipelines (two commercially and one freely available) were used. Most pathogens identified by culture were also identified through metagenomics, but substantial differences were noted between the taxonomic classification tools.

My contribution to this publication included the bioinformatics analysis of all the samples using a unix-based approach. I performed quality assessment and quality control of the WGS and SMg data, the removal of host sequencing from the samples, and the taxonomic identification of the remaining reads in each sample though 3 different methods: MetaPhlan2, Kraken and MIDAS. Gene detection directly from the reads for bacterial typing was also performed using metaMLST, ReMatCh, and Bowtie2 and Samtools. Finally, the reads were assembled using the SPAdes genome assembler, with and without metagenomic mode according to the sample being processed.

@@

Shotgun metagenomics (SMg) is a culture-independent technique that provides valuable information not only at the identification level, but also at the level of molecular characterisation. 
Studies have shown that it has added value in terms of detection sensitivity and personalised treatment in clinical microbiology, when identifying bacteria \citep{hasman_rapid_2014, willmann_antibiotic_2015} or viruses \citep{graf_unbiased_2016}. 
Indeed Gyarmati et al., 2016 \citep{gyarmati_metagenomic_2016}, used a sequence-based metagenomics approach directly from blood to detect non-culturable, difficult-to-culture and non-bacterial pathogens. 
The authors were able, through SMg, to detect viral and fungal pathogens together with bacteria, which had not been detected through classical microbiology. 
Additionally, SMg can be used for infection prevention, having the potential to identify transmission events directly from clinical samples \citep{olson_metagenomic_2017}. 
For example, SMg was proven valuable for the identification of inter-host nucleotide variations occurring after direct transmission of noroviruses causing gastroenteritis \citep{olson_metagenomic_2017}. 
Hasman and colleagues (2014) \citep{hasman_rapid_2014} were able to identify urinary pathogens directly from urine, as well as antimicrobial resistant genes compatible with the resistant phenotype determined through antimicrobial susceptibility testing. They also identified almost perfect phylogenetic matches between whole-genome sequence (WGS) data obtained by metagenomics and WGS of pure isolates.

Despite the promise of SMg of becoming a one-stop solution in clinical microbiology, SMg still has several challenges to overcome. 
One of the greatest challenges is the choice of the extraction and sequencing protocols, as well of the type of controls \citep{schlaberg_validation_2017}. 
The extraction protocol should efficiently and specifically isolate microbial DNA/RNA, while removing the host DNA/RNA \citep{street_molecular_2017}. 
However, the variety of clinical samples used in the diagnosis of distinct types of infection (e.g. tissues versus fluids), poses a serious challenge for standardisation, an essential step if these methods are to be used by routine diagnostic laboratories. 
The sequencing protocol is also dependent on the pathogens of interest (e.g. bacteria versus viruses), sequencing strategy (DNA and/or RNA), required turnaround time, sequencing depth and error tolerance \citep{schlaberg_validation_2017}. The use of defined controls is necessary for validation of each experiment and these should be adapted for every type of infection and sample type and should consist of a combination of known positive specimens, pathogen-negative patient specimens and pathogen-negative patient specimens spiked with live microorganisms or pure DNA \citep{schlaberg_validation_2017}.

Another potential challenge are the metagenomics analysis tools. 
Recent studies have evaluated the different SMg sequence classification methods \citep{peabody_evaluation_2015}. 
These use different methodologies for classification: sequence similarity-based methods, sequence composition-based methods and hybrid methods \citep{peabody_evaluation_2015}. 
They differ not only in the algorithms for detecting the microorganisms present, but also in the databases used. 
This high variability leads to different results, not only at the microorganism classification level but also when evaluating the relative abundance of these pathogens \citep{peabody_evaluation_2015}. 
A recent study evaluated the accuracy of 38 bioinformatics methods using both \textit{in silico} and \textit{in vitro} generated mock bacterial communities.
Dozens to hundreds of species were falsely predicted by the most popular software, and no software clearly outperformed the others \citep{peabody_evaluation_2015}. 
In the absence of studies comparing the outputs of different analysis methods in clinical samples, users may decide which methods to use based on personal experience with a given tool, availability of the tool in the laboratory or its ease of use.
This poses a great challenge when providing reproducible results and creates uncertainty regarding the reliability of the information derived. This is a major barrier to the implementation of SMg approaches in routine clinical microbiology laboratories.

In this study, the aim was to identify the critical steps when using SMg for the identification and characterization of microbial pathogens directly from clinical specimens using methods that are likely to be available in clinical microbiology laboratories wanting to implement genomics for pathogen identification or molecular epidemiology studies. 
For this purpose, we used three human-DNA depletion kits and evaluated a diverse set of bioinformatics tools (commercial and non-commercial) in order to investigate how well they performed and what would the differences be in terms of taxonomic classification, antimicrobial resistance gene detection and typing directly from patient samples, bypassing culture.

@@

In order to evaluate and compare the accuracy and reliability of the bioinformatics analyses in providing the closest results to culture and WGS of any cultured isolates, three different pipelines (two commercially and one freely available) were used (Figure \ref{fig:chap2_figure1}).
Different tools to perform raw read quality control, filtering and trimming were used and reads were mapped against the human genome (hg19) before performing taxonomic classification. 
Reads mapping to hg19 where removed from the analysis to increase the efficiency of the bioinformatics tools. 
Typing (MLST), phylogenetic analysis, plasmid analysis, detection of antimicrobial resistance and virulence genes was performed. 
To determine the appropriateness of SMg as predictor of the WGS (chromosome and plasmids), SMg results obtained were compared with the results of WGS of any bacterial isolates obtained from culturing the sample.

@@

For the WGS reads of the bacterial isolates, the whole INNUca v2.6 pipeline was run, including SPAdes assembly and polishing. 
Plasmids were detected as mentioned previously.

@@

The fastq files containing the reads were uploaded into CLC Genomics Workbench v10.1.1, using the following options: Illumina import, paired-reads, paired-end (forward-reverse) and minimum distance of 1 and a maximum distance of 1000 (default). 
The trimming was performed using the default settings, except the quality trimming score limit was set to 0.01 and we added a Trim adapter list containing Illumina adapters. 
The mapping was performed with the Map Reads to Reference tool, using the hg19 genome as reference. 
The default settings were used with the addition of the collect un-mapped reads option. The \textit{de novo} assembly tool was used for the assembly (even for the metagenomics reads) and, apart from the word size, which was changed to 29, all the settings were default. 
Two tools were used for the microbial identification, Taxonomic Profiling and Find Best Matches using K-mer Spectra (Microbial Genomics Module). 
In both, the bacterial and fungal databases were downloaded from NCBI RefSeq (with the Only Complete Genomes option turned off; minimum length 500,000 nucleotides) on 08-07-2017 (bacterial, 70,868 sequences) and 25-05-2017 (fungal, 377 sequences). 
The antimicrobial resistance genes were detected, based on the assembled contigs, using the Find Resistance tool (Microbial Genomics Module) and were initially only considered present when they were $\geq$ 70\% identical to the reference and $\geq$ 80\% of the sequence was covered. The analysis was also repeated using $\geq$ 40\% and $\geq$ 20\% of sequence coverage for comparison purposes. 
The database containing the antimicrobial resistance genes was downloaded directly to the software from ResFinder\footnote{\url{https://cge.cbs.dtu.dk/services/data.php}} (downloaded on 05-07-2017, 2156 sequences). The MLST was determined through the Identify MLST tool (Microbial Genomics Module), using all MLST schemes available at PubMLST (04-03-2017). 
The same database used for plasmid detection in Unix, was used for mapping the reads in CLC Genomics Workbench. Again, plasmids with less than 80\% of its length covered were excluded from the final results. 
For WGS reads we used the Trim Sequences tool and the assembly, antimicrobial resistance genes detection, and MLST determination were performed as before.

@@

Typing was done by MLST and wgMLST analyses obtained using Ridom SeqSphere+ v4.0.1. 
The genomic data (assembled contigs) obtained from SMg was compared to the data obtained through WGS.
Since no cg/wgMLST scheme was available for \textit{Escherichia coli}, \textit{Enterococcus faecalis}, \textit{Ochrobactrum intermedium} and \textit{Staphylococcus haemolyticus}, cgMLST and accessory genome schemes were constructed, using Ridom SeqSphere+ cgMLST Target Definer with the following parameters: a minimum length filter that removes all genes smaller than 50 bp; a start codon filter that discards all genes that contain no start codon at the beginning of the gene; a stop codon filter that discards all genes that contain no stop codon or more than one stop codon or that do not have the stop codon at the end of the gene; a homologous gene filter that discards all genes with fragments that occur in multiple copies within a genome (with identity of 90\% and >100 bp overlap); and a gene overlap filter that discards the shorter gene from the cgMLST scheme if the two genes affected overlap $>$4 bp. 
The remaining genes were then used in a pairwise comparison using BLAST version 2.2.12 (parameters used were word size 11, mismatch penalty -1, match reward 1, gap open costs 5, and gap extension costs 2). 
All genes of the reference genome that were common in all query genomes with a sequence identity of $\geq$ 90\% and 100\% overlap and, with the default parameter stop codon percentage filter turned on, formed the final cgMLST scheme. 
The combination of all alleles in each strain formed an allelic profile that was used to generate minimum spanning trees using the parameter “pairwise ignore missing values” during distance calculation \citep{ruppitsch_defining_2015}.

@@

\caption{Microorganisms identified by conventional methods, WGS and using shotgun metagenomics and the taxonomic classification methods in Unix.}

@@

  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{WGS-based identification}}} &

@@

\item $^\#$Although there was a laboratory identification, no isolates were available for WGS

@@

\caption{Microorganisms identified by conventional methods, WGS and using shotgun metagenomics and the taxonomic classification methods in CLC Genomics Workbench.}

@@

\caption{Microorganisms identified by conventional methods, WGS and using shotgun metagenomics and the taxonomic classification methods in webpages (BaseSpace, Taxonomer and CosmosID).}

@@

  \multirow{2}{*}{\textbf{WGS-based identification}} &

@@

Metagenomics provides other sequence information in addition to pathogen detection. 
We determined the presence of antimicrobial-resistance genes in the SMg sequence data and compared the results with those obtained from WGS and phenotypic resistance testing (Table \ref{tab:ch2_table6}).

@@

\begin{table}[]
\caption{Antimicrobial resistance phenotypes and antimicrobial resistance genes detected using different approaches.}
\label{tab:ch2_table6}
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|l|l|l|ll|}
\hline
\multicolumn{1}{|c|}{\multirow{2}{*}{\textbf{Sample number}}} &
  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}c@{}}Conventional identification\\  (MALDI-TOF)\end{tabular}}}} &
  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}c@{}}Conventional susceptibility testing\\  (VITEK 2)$^b$\end{tabular}}}} &
  \multicolumn{1}{c|}{\textbf{WGS}} &
  \multicolumn{2}{c|}{\textbf{Shotgun metagenomics}} \\ \cline{4-6} 
\multicolumn{1}{|c|}{} &
  \multicolumn{1}{c|}{} &
  \multicolumn{1}{c|}{} &
  \multicolumn{1}{c|}{\textbf{CLC Genomics Workbench}} &
  \multicolumn{1}{c|}{\textbf{ReMatCh (Unix)}} &
  \multicolumn{1}{c|}{\textbf{CLC Genomics Workbencha}} \\ \hline
1 &
  \begin{tabular}[c]{@{}l@{}}E. faecium\\  S. haemolyticus\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}LEV, ERY, CLI\\  OXA, GEN, CIP, FOS, ERY, CLI\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}erm(B), msr(C), ant(6’)-Ia, aph(3’)-III, dfrG\\  blaZ, mecA, ant(6’)-Ia, aph(3’)-III, \\ aac(6’)-aph(2’’), erm(C), mph(C), msr(A), dfrG\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}erm(B), msr(C), ant(6’)-Ia, aph(3’)-III, \\ aac(6’)-aph(2’’), blaZ, mecA, erm(C), \\ mph(C), msr(A), dfrG\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}erm(B), msr(C), ant(6’)-Ia, aph(3’)-III, \\ aac(6’)-aph(2’’), blaZ, mecA, erm(C), \\ mph(C), msr(A), dfrG\end{tabular} \\ \hline
2 &
  \begin{tabular}[c]{@{}l@{}}E. avium\\  E. coli\\  Anaerobes\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}DOX, CLI\\  susceptible\\  -\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}-\#\\  -\#\\  -\#\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not detected\\  Not detected\\  catS, lnu(D), lsa(C), cepA-44, tet(Q)\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}Not detected\\  Not detected\\  catS, lnu(D), lsa(C), cepA-44, tet(Q), fusA\end{tabular} \\ \hline
3 &
  S. epidermidis &
  OXA, GEN, TEC, FUS, CIP, ERY, CLI &
  -\# &
  \multicolumn{1}{l|}{Not detected} &
  Not detected \\ \hline
4 &
  S. aureus &
  PEN, ERY &
  blaZ, spc, erm(A) &
  \multicolumn{1}{l|}{Not detected} &
  Not detected \\ \hline
5 &
  \begin{tabular}[c]{@{}l@{}}E. coli\\  K. oxytoca\\  S. anginosus\\  E. faecalis\\  Anaerobes\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}susceptible\\  AMX\\  susceptible\\  DOX, CLI\\  -\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}-\#\\  blaOXY-1-3\\  -\#\\  tet(M), lsa(A)\\  -\#\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}-\\  Not detected\\  -\\  tet(M)\\  cfxA4, tet(Q)\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}-\\  Not detected\\  -\\  tet(O)\\  cfxA4, tet(Q)\end{tabular} \\ \hline
6 &
  E. faecium &
  \begin{tabular}[c]{@{}l@{}}PEN, AMX, CFX, IMP, GENhl, \\ STRhl, LEV, ERY, CLI, AMP/SUL\end{tabular} &
  erm(B), msr(C), ant(6’)-Ia, aph(3’)-III, aac(6’)-aph(2’’), dfrG &
  \multicolumn{1}{l|}{Not detected} &
  Not detected \\ \hline
7 &
  S. aureus &
  PEN &
  blaZ &
  \multicolumn{1}{l|}{blaZ, norA} &
  blaZ \\ \hline
8 &
  O. intermedium &
  \begin{tabular}[c]{@{}l@{}}AMX, PIP/TAZ, CFX, CFT, CTZ, \\ IMP, FOX, TOB, FOS, NIT, TMP\end{tabular} &
  blaOCH-2 &
  \multicolumn{1}{l|}{blaOCH-5} &
  blaOCH-2 \\ \hline
9 &
  S. aureus &
  PEN &
  -\# &
  \multicolumn{1}{l|}{blaZ} &
  blaZ \\ \hline
10 &
  S. marcescens &
  AMX, AMC, CFX, FOX, NIT, POL &
  -\# &
  \multicolumn{1}{l|}{blaSST-1, tet(41), oqxB, aac(6’)-Ic} &
  tet(41), oqxB, aac(6’)-Ic \\ \hline
\end{tabular}%
}

@@

In three cases when SMg data covered $\geq$ 93\% of the genome we were able to identify the ST, which corresponded to the one found using WGS of the isolated bacteria using CLC Genomics Workbench (n=2) and metaMLST (n=1). 
These results are summarized in Table \ref{tab:ch2_table7}.
Assembled genomes and metagenomes, were compared by wgMLST analysis using Ridom SeqSphere+. 
Figure \ref{fig:chap2_figure2} shows examples of the allele difference between the genomes obtained through WGS versus the genomes obtained through shotgun metagenomics.

>>>>

As mentioned in Chapter \ref{sssec:shotgun_metagenomics}, \ac{SMg} approaches have been a growing interest to deliver clinically relevant results without \textit{a priori} knowledge of what to expect from a particular clinical sample or patient. 
The capacity to detect all potential pathogens in a sample has great potential utility in the diagnosis of infectious disease. 
However, it is unclear how the variety of available methods impacts the end results.

In this publication \ac{SMg} was applied to nine body fluid samples and one tissue sample from patients at the University Medical Center Groningen (UMCG) with varying degrees of contamination: one sample from peritoneal fluid, five from pus , two from synovial fluid of knees with prosthesis, one from sputum and one from a bone biopsy. The results of microbial identification through whole genome sequencing (\ac{WGS}) and \ac{SMg} were compared to standard culture-based microbiological methods.

In order to evaluate and compare the accuracy and reliability of the bioinformatics analyses in providing the closest results to culture and \ac{WGS} of any cultured isolates, three different bioinformatic pipelines (two commercially and one freely available) were used. Most pathogens identified by culture were also identified through metagenomics, but substantial differences were noted between the taxonomic classification tools.

My contribution to this publication included the bioinformatics analysis of all the samples using a unix-based approach. I performed quality assessment and quality control of the \ac{WGS} and \ac{SMg} data, the removal of host sequencing from the samples, and the taxonomic identification of the remaining reads in each sample though 3 different methods: MetaPhlan2, Kraken and MIDAS. Gene detection directly from the reads for bacterial typing was also performed using metaMLST, ReMatCh, and Bowtie2 and Samtools. Finally, the reads were assembled using the SPAdes genome assembler, with and without metagenomic mode according to the sample being processed.

@@

\ac{SMg} is a culture-independent technique that provides valuable information not only at the identification level, but also at the level of molecular characterisation. 
Studies have shown that it has added value in terms of detection sensitivity and personalised treatment in clinical microbiology, when identifying bacteria \citep{hasman_rapid_2014, willmann_antibiotic_2015} or viruses \citep{graf_unbiased_2016}. 
Indeed Gyarmati et al., 2016 \citep{gyarmati_metagenomic_2016}, used a sequence-based metagenomics approach directly from blood to detect non-culturable, difficult-to-culture and non-bacterial pathogens. 
The authors were able, through \ac{SMg}, to detect viral and fungal pathogens together with bacteria, which had not been detected through classical microbiology. 
Additionally, \ac{SMg} can be used for infection prevention, having the potential to identify transmission events directly from clinical samples \citep{olson_metagenomic_2017}. 
For example, \ac{SMg} was proven valuable for the identification of inter-host nucleotide variations occurring after direct transmission of noroviruses causing gastroenteritis \citep{olson_metagenomic_2017}. 
Hasman and colleagues (2014) \citep{hasman_rapid_2014} were able to identify urinary pathogens directly from urine, as well as antimicrobial resistant genes compatible with the resistant phenotype determined through antimicrobial susceptibility testing. They also identified almost perfect phylogenetic matches between \ac{WGS} data obtained by metagenomics and \ac{WGS} of pure isolates.

Despite the promise of \ac{SMg} of becoming a one-stop solution in clinical microbiology, \ac{SMg} still has several challenges to overcome. 
One of the greatest challenges is the choice of the extraction and sequencing protocols, as well of the type of controls \citep{schlaberg_validation_2017}. 
The extraction protocol should efficiently and specifically isolate microbial DNA/RNA, while removing the host DNA/RNA \citep{street_molecular_2017}. 
However, the variety of clinical samples used in the diagnosis of distinct types of infection (e.g. tissues versus fluids), poses a serious challenge for standardisation, an essential step if these methods are to be used by routine diagnostic laboratories. 
The sequencing protocol is also dependent on the pathogens of interest (e.g. bacteria versus viruses), sequencing strategy (DNA and/or RNA), required turnaround time, sequencing depth and error tolerance \citep{schlaberg_validation_2017}. The use of defined controls is necessary for validation of each experiment and these should be adapted for every type of infection and sample type and should consist of a combination of known positive specimens, pathogen-negative patient specimens and pathogen-negative patient specimens spiked with live microorganisms or pure DNA \citep{schlaberg_validation_2017}.

Another potential challenge are the metagenomics analysis tools. 
Recent studies have evaluated the different \ac{SMg} sequence classification methods \citep{peabody_evaluation_2015}. 
These use different methodologies for classification: sequence similarity-based methods, sequence composition-based methods and hybrid methods \citep{peabody_evaluation_2015}. 
They differ not only in the algorithms for detecting the microorganisms present, but also in the databases used. 
This high variability leads to different results, not only at the microorganism classification level but also when evaluating the relative abundance of these pathogens \citep{peabody_evaluation_2015}. 
A recent study evaluated the accuracy of 38 bioinformatics methods using both \textit{in silico} and \textit{in vitro} generated mock bacterial communities.
Dozens to hundreds of species were falsely predicted by the most popular software, and no software clearly outperformed the others \citep{peabody_evaluation_2015}. 
In the absence of studies comparing the outputs of different analysis methods in clinical samples, users may decide which methods to use based on personal experience with a given tool, availability of the tool in the laboratory or its ease of use.
This poses a great challenge when providing reproducible results and creates uncertainty regarding the reliability of the information derived. This is a major barrier to the implementation of \ac{SMg} approaches in routine clinical microbiology laboratories.

In this study, the aim was to identify the critical steps when using \ac{SMg} for the identification and characterization of microbial pathogens directly from clinical specimens using methods that are likely to be available in clinical microbiology laboratories wanting to implement genomics for pathogen identification or molecular epidemiology studies. 
For this purpose, we used three human-DNA depletion kits and evaluated a diverse set of bioinformatics tools (commercial and non-commercial) in order to investigate how well they performed and what would the differences be in terms of taxonomic classification, antimicrobial resistance gene detection and typing directly from patient samples, bypassing culture.

@@

In order to evaluate and compare the accuracy and reliability of the bioinformatics analyses in providing the closest results to culture and \ac{WGS} of any cultured isolates, three different pipelines (two commercially and one freely available) were used (Figure \ref{fig:chap2_figure1}).
Different tools to perform raw read quality control, filtering and trimming were used and reads were mapped against the human genome (hg19) before performing taxonomic classification. 
Reads mapping to hg19 where removed from the analysis to increase the efficiency of the bioinformatics tools. 
Typing (MLST), phylogenetic analysis, plasmid analysis, detection of antimicrobial resistance and virulence genes was performed. 
To determine the appropriateness of \ac{SMg} as predictor of the \ac{WGS} (chromosome and plasmids), \ac{SMg} results obtained were compared with the results of \ac{WGS} of any bacterial isolates obtained from culturing the sample.

@@

For the \ac{WGS} reads of the bacterial isolates, the whole INNUca v2.6 pipeline was run, including SPAdes assembly and polishing. 
Plasmids were detected as mentioned previously.

@@

The fastq files containing the reads were uploaded into CLC Genomics Workbench v10.1.1, using the following options: Illumina import, paired-reads, paired-end (forward-reverse) and minimum distance of 1 and a maximum distance of 1000 (default). 
The trimming was performed using the default settings, except the quality trimming score limit was set to 0.01 and we added a Trim adapter list containing Illumina adapters. 
The mapping was performed with the Map Reads to Reference tool, using the hg19 genome as reference. 
The default settings were used with the addition of the collect un-mapped reads option. The \textit{de novo} assembly tool was used for the assembly (even for the metagenomics reads) and, apart from the word size, which was changed to 29, all the settings were default. 
Two tools were used for the microbial identification, Taxonomic Profiling and Find Best Matches using K-mer Spectra (Microbial Genomics Module). 
In both, the bacterial and fungal databases were downloaded from NCBI RefSeq (with the Only Complete Genomes option turned off; minimum length 500,000 nucleotides) on 08-07-2017 (bacterial, 70,868 sequences) and 25-05-2017 (fungal, 377 sequences). 
The antimicrobial resistance genes were detected, based on the assembled contigs, using the Find Resistance tool (Microbial Genomics Module) and were initially only considered present when they were $\geq$ 70\% identical to the reference and $\geq$ 80\% of the sequence was covered. The analysis was also repeated using $\geq$ 40\% and $\geq$ 20\% of sequence coverage for comparison purposes. 
The database containing the antimicrobial resistance genes was downloaded directly to the software from ResFinder\footnote{\url{https://cge.cbs.dtu.dk/services/data.php}} (downloaded on 05-07-2017, 2156 sequences). The MLST was determined through the Identify MLST tool (Microbial Genomics Module), using all MLST schemes available at PubMLST (04-03-2017). 
The same database used for plasmid detection in Unix, was used for mapping the reads in CLC Genomics Workbench. Again, plasmids with less than 80\% of its length covered were excluded from the final results. 
For \ac{WGS} reads we used the Trim Sequences tool and the assembly, antimicrobial resistance genes detection, and MLST determination were performed as before.

@@

Typing was done by MLST and wgMLST analyses obtained using Ridom SeqSphere+ v4.0.1. 
The genomic data (assembled contigs) obtained from \ac{SMg} was compared to the data obtained through \ac{WGS}.
Since no cg/wg MLST scheme was available for \textit{Escherichia coli}, \textit{Enterococcus faecalis}, \textit{Ochrobactrum intermedium} and \textit{Staphylococcus haemolyticus}, cgMLST and accessory genome schemes were constructed, using Ridom SeqSphere+ cgMLST Target Definer with the following parameters: a minimum length filter that removes all genes smaller than 50 bp; a start codon filter that discards all genes that contain no start codon at the beginning of the gene; a stop codon filter that discards all genes that contain no stop codon or more than one stop codon or that do not have the stop codon at the end of the gene; a homologous gene filter that discards all genes with fragments that occur in multiple copies within a genome (with identity of 90\% and >100 bp overlap); and a gene overlap filter that discards the shorter gene from the cgMLST scheme if the two genes affected overlap $>$4 bp. 
The remaining genes were then used in a pairwise comparison using BLAST version 2.2.12 (parameters used were word size 11, mismatch penalty -1, match reward 1, gap open costs 5, and gap extension costs 2). 
All genes of the reference genome that were common in all query genomes with a sequence identity of $\geq$ 90\% and 100\% overlap and, with the default parameter stop codon percentage filter turned on, formed the final cgMLST scheme. 
The combination of all alleles in each strain formed an allelic profile that was used to generate minimum spanning trees using the parameter “pairwise ignore missing values” during distance calculation \citep{ruppitsch_defining_2015}.

@@

\caption{Microorganisms identified by conventional methods, \ac{WGS} and using shotgun metagenomics and the taxonomic classification methods in Unix.}

@@

  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{\ac{WGS}-based identification}}} &

@@

\item $^\#$Although there was a laboratory identification, no isolates were available for \ac{WGS}

@@

\caption{Microorganisms identified by conventional methods, \ac{WGS} and using shotgun metagenomics and the taxonomic classification methods in CLC Genomics Workbench.}

@@

\caption{Microorganisms identified by conventional methods, \ac{WGS} and using shotgun metagenomics and the taxonomic classification methods in webpages (BaseSpace, Taxonomer and CosmosID).}

@@

  \multirow{2}{*}{\textbf{\ac{WGS}-based identification}} &

@@

Metagenomics provides other sequence information in addition to pathogen detection. 
We determined the presence of antimicrobial-resistance genes in the \ac{SMg} sequence data and compared the results with those obtained from \ac{WGS} and phenotypic resistance testing (Table \ref{tab:ch2_table6}).

@@

\begin{table}[]
\caption{Antimicrobial resistance phenotypes and antimicrobial resistance genes detected using different approaches.}
\label{tab:ch2_table6}
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|l|l|l|ll|}
\hline
\multicolumn{1}{|c|}{\multirow{2}{*}{\textbf{Sample number}}} &
  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}c@{}}Conventional identification\\  (MALDI-TOF)\end{tabular}}}} &
  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}c@{}}Conventional susceptibility testing\\  (VITEK 2)$^b$\end{tabular}}}} &
  \multicolumn{1}{c|}{\textbf{\ac{WGS}}} &
  \multicolumn{2}{c|}{\textbf{Shotgun metagenomics}} \\ \cline{4-6} 
\multicolumn{1}{|c|}{} &
  \multicolumn{1}{c|}{} &
  \multicolumn{1}{c|}{} &
  \multicolumn{1}{c|}{\textbf{CLC Genomics Workbench}} &
  \multicolumn{1}{c|}{\textbf{ReMatCh (Unix)}} &
  \multicolumn{1}{c|}{\textbf{CLC Genomics Workbencha}} \\ \hline
1 &
  \begin{tabular}[c]{@{}l@{}}E. faecium\\  S. haemolyticus\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}LEV, ERY, CLI\\  OXA, GEN, CIP, FOS, ERY, CLI\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}erm(B), msr(C), ant(6’)-Ia, aph(3’)-III, dfrG\\  blaZ, mecA, ant(6’)-Ia, aph(3’)-III, \\ aac(6’)-aph(2’’), erm(C), mph(C), msr(A), dfrG\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}erm(B), msr(C), ant(6’)-Ia, aph(3’)-III, \\ aac(6’)-aph(2’’), blaZ, mecA, erm(C), \\ mph(C), msr(A), dfrG\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}erm(B), msr(C), ant(6’)-Ia, aph(3’)-III, \\ aac(6’)-aph(2’’), blaZ, mecA, erm(C), \\ mph(C), msr(A), dfrG\end{tabular} \\ \hline
2 &
  \begin{tabular}[c]{@{}l@{}}E. avium\\  E. coli\\  Anaerobes\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}DOX, CLI\\  susceptible\\  -\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}-\#\\  -\#\\  -\#\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not detected\\  Not detected\\  catS, lnu(D), lsa(C), cepA-44, tet(Q)\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}Not detected\\  Not detected\\  catS, lnu(D), lsa(C), cepA-44, tet(Q), fusA\end{tabular} \\ \hline
3 &
  S. epidermidis &
  OXA, GEN, TEC, FUS, CIP, ERY, CLI &
  -\# &
  \multicolumn{1}{l|}{Not detected} &
  Not detected \\ \hline
4 &
  S. aureus &
  PEN, ERY &
  blaZ, spc, erm(A) &
  \multicolumn{1}{l|}{Not detected} &
  Not detected \\ \hline
5 &
  \begin{tabular}[c]{@{}l@{}}E. coli\\  K. oxytoca\\  S. anginosus\\  E. faecalis\\  Anaerobes\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}susceptible\\  AMX\\  susceptible\\  DOX, CLI\\  -\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}-\#\\  blaOXY-1-3\\  -\#\\  tet(M), lsa(A)\\  -\#\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}-\\  Not detected\\  -\\  tet(M)\\  cfxA4, tet(Q)\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}-\\  Not detected\\  -\\  tet(O)\\  cfxA4, tet(Q)\end{tabular} \\ \hline
6 &
  E. faecium &
  \begin{tabular}[c]{@{}l@{}}PEN, AMX, CFX, IMP, GENhl, \\ STRhl, LEV, ERY, CLI, AMP/SUL\end{tabular} &
  erm(B), msr(C), ant(6’)-Ia, aph(3’)-III, aac(6’)-aph(2’’), dfrG &
  \multicolumn{1}{l|}{Not detected} &
  Not detected \\ \hline
7 &
  S. aureus &
  PEN &
  blaZ &
  \multicolumn{1}{l|}{blaZ, norA} &
  blaZ \\ \hline
8 &
  O. intermedium &
  \begin{tabular}[c]{@{}l@{}}AMX, PIP/TAZ, CFX, CFT, CTZ, \\ IMP, FOX, TOB, FOS, NIT, TMP\end{tabular} &
  blaOCH-2 &
  \multicolumn{1}{l|}{blaOCH-5} &
  blaOCH-2 \\ \hline
9 &
  S. aureus &
  PEN &
  -\# &
  \multicolumn{1}{l|}{blaZ} &
  blaZ \\ \hline
10 &
  S. marcescens &
  AMX, AMC, CFX, FOX, NIT, POL &
  -\# &
  \multicolumn{1}{l|}{blaSST-1, tet(41), oqxB, aac(6’)-Ic} &
  tet(41), oqxB, aac(6’)-Ic \\ \hline
\end{tabular}%
}

@@

In three cases when \ac{SMg} data covered $\geq$ 93\% of the genome we were able to identify the ST, which corresponded to the one found using \ac{WGS} of the isolated bacteria using CLC Genomics Workbench (n=2) and metaMLST (n=1). 
These results are summarized in Table \ref{tab:ch2_table7}.
Assembled genomes and metagenomes, were compared by wgMLST analysis using Ridom SeqSphere+. 
Figure \ref{fig:chap2_figure2} shows examples of the allele difference between the genomes obtained through \ac{WGS} versus the genomes obtained through shotgun metagenomics.
